---
title: "Tables for Manuscript"
author: "Simran Makwana & Alba Gutiérrez"
date: "2023-10-16"
output:
  word_document: default
  html_document: default
---

## Load the required libraries
```{r}
library(table1)
library(dplyr)
library(tidyr)
library(stringr)
```

##Step 1: Run reformat_aggregate_output.Rmd with latest outputs to get combined dataframes with all data. Load the two files (a1 and a2 RData files)

```{r}
load(file = '../results/bacterial_aggregate_ouputs/a1_all_to_work.RData')
load(file = '../results/bacterial_aggregate_ouputs/a2_all_to_work.RData')
```

Capitalize the first letter to each category name for harmonization.
```{r}
a2_all_to_work$category <- stringr::str_to_title( a2_all_to_work$category )
a1_all_to_work$category <- stringr::str_to_title( a1_all_to_work$category )
```

Organize sites into countries
- This list must be updated if additional countries / sites are added
```{r}
site_country_mapping <- data.frame('site' = c('BCH', 'CCHMC', 'CCMC', 'GOSH', 'H12O', 'HCUV', 'MICHIGAN', 'UPitt'),
                                   'country' = c('United States', 'United States', 'United States', 'United Kingdom', 
                                                 'Spain', 'Spain', 'United States', 'United States'))

a1.1 <- a1_all_to_work %>%
  mutate(site = gsub('_.*', '', site), 
         site = trimws(site)) %>%
  left_join(site_country_mapping)

a2.1 <- a2_all_to_work %>%
  mutate(site = gsub('_.*', '', site),
         site = trimws(site)) %>%
  left_join(site_country_mapping)

rm(a1_all_to_work)
rm(a2_all_to_work)
```
# Count function
We need to estimate the number and percentage of hospitalizations, by period (pre-NPI, full-NPI, partial-NPI), based on different scenarios:
- filtering by type == all when we want table 1
- filtering by type == bacterial infection when we want table X
- doing the counts at different levels (age, sex, etc)

We create a function that given the type and variable generates the table. Then we apply it to the different conditions. 
- Percentages are calculated by clinical characteristic (columnwise)... for example all USA, UK and Spain counts within the pre-npi cohort add to 100%. To change this, modify `group_by(timeToGroup)` to something like `group_by(country)` in each section.

```{r, echo=FALSE}
countsAndPercentages <- function( input_df, selected_type, variable, percentage ){
  
  if( variable == "country"){
    output <- input_df %>%
      dplyr::filter(type == selected_type) %>%
      dplyr::filter(category_name == "total" ) 
      
    output <- output %>%
      dplyr::group_by(timeToGroup, country) %>%
      dplyr::summarise(n = sum(count)) %>%
      dplyr::ungroup() %>%
      dplyr::group_by(timeToGroup) %>%
      dplyr::mutate(percent = n / sum(n) * 100,
                    value = paste0(n, ' (', round(percent, 2), '%)')) 
    
    output_wider <- output %>%
      tidyr::pivot_wider(id_cols = country, 
                         names_from = timeToGroup, 
                         values_from = value) %>%
      dplyr::add_row(country = 'Country, n (%)', .before = 1) %>%
      dplyr::rename('X' = 'country')
  }else if( variable == "subgroup"){
    output <- input_df %>%
      dplyr::filter(type == selected_type) %>%
      dplyr::filter(category_name == "subgroups" ) 
      
    output <- output %>%
      dplyr::group_by(timeToGroup, subgroup) %>%
      dplyr::summarise(n = sum(count)) %>%
      dplyr::ungroup() 
    
    if( percentage == TRUE){
      output <- output %>%
        dplyr::group_by(timeToGroup) %>%
        dplyr::mutate(percent = n / sum(n) * 100,
                    value = paste0(n, ' (', round(percent, 2), '%)'))
    }else{
       output <- output %>%
        dplyr::group_by(timeToGroup) %>%
        dplyr::mutate(value = n)
    }
    
    output_wider <- output %>%
      tidyr::pivot_wider(id_cols = subgroup, 
                         names_from = timeToGroup, 
                         values_from = value)

  }else{
    output <- input_df %>%
      dplyr::filter(type == selected_type) %>%
      dplyr::filter(category_name == variable ) 
      
    output <- output %>%
      dplyr::group_by(timeToGroup, category) %>%
      dplyr::summarise(n = sum(count)) %>%
      dplyr::ungroup()
    
    if( percentage == TRUE){
      output <- output %>%
        dplyr::group_by(timeToGroup) %>%
        dplyr::mutate(percent = n / sum(n) * 100,
                    value = paste0(n, ' (', round(percent, 2), '%)'))
    }else{
       output <- output %>%
        dplyr::group_by(timeToGroup) %>%
        dplyr::mutate(value = n)
    }
    
    output_wider <- output %>%
      tidyr::pivot_wider(id_cols = category, 
                         names_from = timeToGroup, 
                         values_from = value) %>%
      dplyr::add_row(category = paste0( variable, ', n (%)'), .before = 1) %>%
      dplyr::rename('X' = 'category')
  }
  
  return( output_wider )
}
```



# Table 1 
- Filter to type == all; include all study population hospitalizations
- *NPI cohorts are redefined for the season counts using the a2 tables, verify that the cutoffs are correct*
```{r echo = FALSE}
# Country
t1.country <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "all", 
                                 variable = "country", 
                                 percentage = TRUE)
# Age at hospital admission
t1.age <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "all", 
                                 variable = "age", 
                                 percentage = TRUE)
orderAge <- c("age, n (%)","Infancy", "Childhood", "Middle Childhood", "Adolescence")
t1.age$X <- factor(t1.age$X , levels = orderAge)
t1.age <- t1.age[order(t1.age$X), ]


# Sex
t1.sex <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "all", 
                                 variable = "sex", 
                                 percentage = TRUE)
# Season
t1.season <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "all", 
                                 variable = "season", 
                                 percentage = TRUE)

orderSeason <- c("season, n (%)","Spring", "Summer", "Autumn", "Winter")
t1.season$X <- factor(t1.season$X , levels = orderSeason)
t1.season <- t1.season[order(t1.season$X), ]


# Lenght of hospitalization
t1.lengthHospitalization <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "all", 
                                 variable = "lengthHospitalization", 
                                 percentage = TRUE)

orderLength <- c("lengthHospitalization, n (%)","Less Than 3 Days", "Between 3 And 7 Days", "More Than 7 Days")
t1.lengthHospitalization$X <- factor(t1.lengthHospitalization$X , levels = orderLength)
t1.lengthHospitalization <- t1.lengthHospitalization[order(t1.lengthHospitalization$X), ]
```

Combine Table 1 sub-tables
```{r echo = FALSE}
table1 <- rbind(t1.country, t1.age, t1.sex, t1.season, t1.lengthHospitalization) %>%
  select(X, `pre-NPI`, `full-NPI`, `partial-NPI`)

colnames(table1) <- c(" ", "pre-NPI \n Jan 2019-Feb 2020", 
                      "full-NPI \n March 2020-Feb 2021", 
                      "partial-NPI \n March 2021-Feb 2023")
```

```{r}
rm(list=setdiff(ls(), c("table1","a1.1", "countsAndPercentages")))

```


# Table 2 

```{r}
table2 <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable ="subgroup", 
                                 percentage = FALSE)

table2 <- table2 %>%
  select(subgroup, `pre-NPI`, `full-NPI`, `partial-NPI`)

```


# Table 3 

```{r}
table3 <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable ="bacterialInfection", 
                                 percentage = FALSE)
table3 <- table3 %>%
  select(`X`, `pre-NPI`, `full-NPI`, `partial-NPI`)
```

# Table X 
- Filter to type == bacterial
```{r echo = FALSE}
# Country
tX.country <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable = "country", 
                                 percentage = TRUE)
# Age at hospital admission
tX.age <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable = "age", 
                                 percentage = TRUE)

orderAge <- c("age, n (%)","infancy", "Childhood", "Middle Childhood", "Adolescence")
tX.age$X <- factor(tX.age$X , levels = orderAge)
tX.age <- tX.age[order(tX.age$X), ]

# Sex
tX.sex <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable = "sex", 
                                 percentage = TRUE)
# Season
tX.season <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable = "season", 
                                 percentage = TRUE)
orderSeason <- c("season, n (%)","Spring", "Summer", "Autumn", "Winter")
tX.season$X <- factor(tX.season$X , levels = orderSeason)
tX.season <- tX.season[order(tX.season$X), ]
# Lenght of hospitalization
tX.lengthHospitalization <- countsAndPercentages( input_df = a1.1, 
                                 selected_type = "bacterial", 
                                 variable = "lengthHospitalization", 
                                 percentage = TRUE)

orderLength <- c("lengthHospitalization, n (%)","Less Than 3 Days", "Between 3 And 7 Days", "More Than 7 Days")
tX.lengthHospitalization$X <- factor(tX.lengthHospitalization$X , levels = orderLength)
tX.lengthHospitalization <- tX.lengthHospitalization[order(tX.lengthHospitalization$X), ]
```

Combine Table X sub-tables
```{r echo = FALSE}
tableX <- rbind(tX.country, tX.age, tX.sex, tX.season, tX.lengthHospitalization) %>%
  select(X, `pre-NPI`, `full-NPI`, `partial-NPI`)

```

```{r}
rm(list=setdiff(ls(), c("table1","table2","table3","tableX","a1.1", "countsAndPercentages")))
```

```{r}
#install.packages("officer")
#install.packages("flextable")
library(officer)
library(flextable)

# Create a Word document
doc <- read_docx()

# Add a heading for table1
doc <- body_add_par(doc, "Table 1", 
                    style = "heading 1")

# Add the table1 to the doc
doc <- body_add_flextable(
  doc,
  value = flextable(table1))

# Add a heading for table2
doc <- body_add_par(doc, "Table 2", 
                    style = "heading 1")

# Add the table2 to the doc
doc <- body_add_flextable(
  doc,
  value = flextable(table2))


# Add a heading for table3
doc <- body_add_par(doc, "Table 3",
                    style = "heading 1")

# Add the table 3 to the doc
doc <- body_add_flextable(
  doc,
  value = flextable(table3))

# Add a heading for tableX
doc <- body_add_par(doc, "Table X",
                    style = "heading 1")

# Add the table X to the doc
doc <- body_add_flextable(
  doc,
  value = flextable(tableX))

# Save the document
print(doc, target = "../../bacterialTables.docx")

```
