---
title: "Tables for Manuscript"
author: "Simran Makwana"
date: "2023-10-13"
output: html_document
---

##Step 1: Run reformat_aggregate_output.Rmd with latest outputs to get combined dataframes with all data.
- The output of that script is loaded here
- Table 1 package is loaded
- **Outputs including RSV are excluded**

```{r echo = FLASE}
library(table1)

load(file = '../results/a1_all.Rdata')
load(file = '../results/a2_all.Rdata')

a1 <- a1_all_to_work %>% 
  filter(!grepl('withRSV', site))

a2 <- a2_all_to_work %>% 
  filter(!grepl('withRSV', site))
```

Organize sites into countries
- This list must be updated if additional countries / sites are added
```{r}
site_country_mapping <- data.frame('site' = c('BCH', 'CCHMC', 'CCMC', 'GOSH', 'H12O', 'HCUV', 'MICHIGAN', 'UPitt'),
                                   'country' = c('United States', 'United States', 'United States', 'United Kingdom', 
                                                 'Spain', 'Spain', 'United States', 'United States'))

a1.1 <- a1 %>%
  mutate(site = gsub('_.*', '', site), 
         site = trimws(site)) %>%
  left_join(site_country_mapping)

a2.1 <- a2 %>%
  mutate(site = gsub('_.*', '', site),
         site = trimws(site)) %>%
  left_join(site_country_mapping)

```

# Table 1:  Clinical characteristics of the study population, stratified by cohort  
- Filter to type == all; include all study population hospitalizations
- Percentages are calculated by clinical characteristic (columnwise)... for example all USA, UK and Spain counts within the pre-npi cohort add to 100%. To change this, modify `group_by(timeToGroup)` to something like `group_by(country)` in each section.
- *NPI cohorts are redefined for the season counts using the a2 tables, verify that the cutoffs are correct*
- *We had some data out of scope, after February 2023. currently removing it*
```{r echo = FALSE}

# Country
t1.country <- a1.1 %>%
  filter(type == 'all') %>%
  filter(category_name == 'total') %>%
  group_by(timeToGroup, country) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = country, names_from = timeToGroup, values_from = value) %>%
  add_row(country = 'Country, n (%)', .before = 1) %>%
  rename('X' = 'country')

# Age at hospital admission
t1.age <- a1.1 %>%
  filter(type == 'all') %>%
  filter(category_name == 'age') %>%
  group_by(timeToGroup, category) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = category, names_from = timeToGroup, values_from = value) %>%
  add_row(category = 'Age at hospital admission, n (%)', .before = 1) %>%
  rename('X' = 'category')

# Sex
t1.sex <- a1.1 %>%
  filter(type == 'all') %>%
  filter(category_name == 'sex') %>%
  mutate(category = tolower(category)) %>%
  group_by(timeToGroup, category) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = category, names_from = timeToGroup, values_from = value) %>%
  add_row(category = 'Sex, n (%)', .before = 1) %>%
  rename('X' = 'category')

# Season
month_season_mapping <- data.frame(month = c(3:12, 1, 2),
                                   season = c(rep('Spring', 3), rep('Summer', 3), rep('Autumn', 3), rep('Winter', 3))) %>%
  mutate(season = factor(season, levels = c('Spring', 'Summer', 'Autumn', 'Winter')))

t1.season <- a2.1 %>% 
  filter(type == 'all') %>%
  mutate(month = month(timeToGroup),
# should check to make sure I defined these cohorts correctly
         cohort = ifelse(as.Date('2019-03-01') <= timeToGroup & timeToGroup < as.Date('2020-02-28'), 'pre-NPI',
                         ifelse(timeToGroup < as.Date('2021-02-28'), 'full-NPI', 
                                ifelse(timeToGroup < as.Date('2023-02-28'), 'partial-NPI',
                                                             'out of scope')))) %>%
  filter(cohort != 'out of scope') %>%
  left_join(month_season_mapping) %>%
  group_by(cohort, season) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(cohort) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = season, names_from = cohort, values_from = value) %>%
  add_row(season = 'Season, n (%)', .before = 1) %>%
  rename('X' = 'season')
  

# Length of hospitalization
# where did we put this? I thought we made groups for the length of hospitalization?
```

Combine Table 1 sub-tables
```{r echo = FALSE}
table1 <- rbind(t1.country, t1.age, t1.sex, t1.season) %>%
  select(X, `pre-NPI`, `full-NPI`, `partial-NPI`)

```


# Table X (Appendix). Clinical characteristics of the hospitalisations with any bacterial infection, stratified by cohort  
- Filter to type == bacterial; include only hospitalizations with bacterial infections
- Percentages are calculated by clinical characteristic (columnwise)... for example all USA, UK and Spain counts within the pre-npi cohort add to 100%. To change this, modify `group_by(timeToGroup)` to something like `group_by(country)` in each section.
- *NPI cohorts are redefined for the season counts using the a2 tables, verify that the cutoffs are correct*
- *We had some data out of scope, after February 2023. currently removing it*
```{r echo = FALSE}

# Country
tX.country <- a1.1 %>%
  filter(type == 'bacterial') %>%
  filter(category_name == 'total') %>%
  group_by(timeToGroup, country) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = country, names_from = timeToGroup, values_from = value) %>%
  add_row(country = 'Country, n (%)', .before = 1) %>%
  rename('X' = 'country')

# Age at hospital admission
tX.age <- a1.1 %>%
  filter(type == 'bacterial') %>%
  filter(category_name == 'age') %>%
  group_by(timeToGroup, category) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = category, names_from = timeToGroup, values_from = value) %>%
  add_row(category = 'Age at hospital admission, n (%)', .before = 1) %>%
  rename('X' = 'category')

# Sex
tX.sex <- a1.1 %>%
  filter(type == 'bacterial') %>%
  filter(category_name == 'sex') %>%
  mutate(category = tolower(category)) %>%
  group_by(timeToGroup, category) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(timeToGroup) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = category, names_from = timeToGroup, values_from = value) %>%
  add_row(category = 'Sex, n (%)', .before = 1) %>%
  rename('X' = 'category')

# Season
month_season_mapping <- data.frame(month = c(3:12, 1, 2),
                                   season = c(rep('Spring', 3), rep('Summer', 3), rep('Autumn', 3), rep('Winter', 3))) %>%
  mutate(season = factor(season, levels = c('Spring', 'Summer', 'Autumn', 'Winter')))

tX.season <- a2.1 %>% 
  filter(type == 'bacterial') %>%
  mutate(month = month(timeToGroup),
# should check to make sure I defined these cohorts correctly
         cohort = ifelse(as.Date('2019-03-01') <= timeToGroup & timeToGroup < as.Date('2020-02-28'), 'pre-NPI',
                         ifelse(timeToGroup < as.Date('2021-02-28'), 'full-NPI', 
                                ifelse(timeToGroup < as.Date('2023-02-28'), 'partial-NPI',
                                                             'out of scope')))) %>%
  filter(cohort != 'out of scope') %>%
  left_join(month_season_mapping) %>%
  group_by(cohort, season) %>%
  summarise(n = sum(count)) %>%
  ungroup() %>%
  group_by(cohort) %>%
  mutate(percent = n / sum(n) * 100,
         value = paste0(n, ' (', round(percent, 2), '%)')) %>%
  pivot_wider(id_cols = season, names_from = cohort, values_from = value) %>%
  add_row(season = 'Season, n (%)', .before = 1) %>%
  rename('X' = 'season')
  

# Length of hospitalization
# where did we put this? I thought we made groups for the length of hospitalization?
```

Combine Table 1 sub-tables
```{r echo = FALSE}
tableX <- rbind(tX.country, tX.age, tX.sex, tX.season) %>%
  select(X, `pre-NPI`, `full-NPI`, `partial-NPI`)

```

