---
title: "4CE Bacterial Infection: reformat aggregate counts output for Analysis"
author: "Alba Gutierrez Sacristan"
date: "2023-08-08"
output: html_document
---

## Overview

For the bacterial infection project, that follows a federated approach, each site is running the "bacterialInfection.Rmd" to generate a set of data frames, that contains the counts at different level of aggregation. 

Here we follow a step-by-step approach to put all those output together, and reformat it to make it easy to use for the meta-analysis. 

## Set up

### Load libraries
```{r }
library(dplyr)
library(lubridate)
```

### Determine the location of the files
The folder containing the outputs from the different sites follows the next structure:
- One folder per site with the name of the site
- Inside each site folder, two .RData files, "a1Tables.RData" and "a2Tables.RData"
```{r}
dir.input <- "../results/QC_counts/"
```

#### Get all the folders names 
```{r}
files <- list.files(dir.input)
```

#### For loop reading each data frame and making some modifications
When reading the different data frames we will add 3 columns:
- category: that can be total, age, sex dead, icu and bacterial infection
- type: that can be all or bacterial infection
- site: to distingish between the different sites

##### A2 tables
```{r}
for( i in 1:length( files )){
  # files for each site
  site_id <- files[i]
  
  #load the RData file with all the a2 tables (aggregate counts per month)
  load(paste0(dir.input, files[i], "/", "a2Tables.RData"))
  
  a2_total_all <-a2_total_all %>%
    dplyr::mutate( category_name = "total", type = "all" )
  
  a2_age_all <- a2_age_all %>%
    dplyr::mutate( category_name = "age", type = "all" )
  
  a2_sex_all <- a2_sex_all %>%
    dplyr::mutate( category_name = "sex", type = "all" )
  
  a2_lengthHosp_all <- a2_lengthHosp_all %>%
    dplyr::mutate( category_name = "lengthHospitalization", type = "all" )
  
  a2_total_bacterial <-a2_total_bacterial %>%
    dplyr::mutate( category_name = "total", type = "bacterial" )
  
  a2_age_bacterial <- a2_age_bacterial %>%
    dplyr::mutate( category_name = "age", type = "bacterial" )
  
  a2_sex_bacterial <- a2_sex_bacterial %>%
    dplyr::mutate( category_name = "sex", type = "bacterial" )
  
  a2_bact_infection <- a2_bact_infection %>%
    dplyr::mutate( category_name = "bacterialInfection", type = "bacterial" ) %>%
    dplyr::rename( category = disorder_group )
  
  a2_lengthHosp_bacterial <- a2_lengthHosp_bacterial %>%
    dplyr::mutate( category_name = "lengthHospitalization", type = "bacterial" ) 
  
   a2_subgroup_bacterial <- a2_subgroups %>%
    dplyr::mutate( category_name = "subgroups", type = "bacterial" )
 
  
  ### add dead and ICU grouped by year
  a2_icu_all <- a2_icu_all %>%
    dplyr::mutate( category_name = "icu",type = "all" )
  
  a2_icu_bacterial <- a2_icu_bacterial %>%
    dplyr::mutate( category_name = "icu", type = "bacterial" )
  
  a2_mortality_all <- a2_mortality_all %>%
    dplyr::mutate( category_name = "dead", type = "all" )
  
  a2_mortality_bacterial <- a2_mortality_bacterial %>%
    dplyr::mutate( category_name = "dead", type = "bacterial" )
  
  a2_icu_mortality <- rbind( a2_icu_all, a2_icu_bacterial,
                             a2_mortality_all, a2_mortality_bacterial)
  
  # remove the original data frames
  rm(a2_icu_all, a2_icu_bacterial, a2_mortality_all, a2_mortality_bacterial)
  
  # rename some columns and put it in the same format to combine with the rest of a2 tables
  a2_icu_mortality <- a2_icu_mortality %>%
    dplyr::mutate( year_val = year( timeToGroup),
                   timeToGroup = lubridate::ymd( year_val, truncated = 2L )) %>%
    dplyr::group_by( timeToGroup, category, category_name, type ) %>%
    dplyr::summarise( count = sum(count )) %>%
    dplyr::mutate( category = as.character( category )) %>%
    dplyr::select( timeToGroup, category, count, category_name, type )
  
  # add all the a2 tables together as a2_all and remove the original ones 
  # add a column with the site id
  a2_all <- rbind( a2_age_all, a2_sex_all, a2_total_all, 
                   a2_lengthHosp_all, a2_lengthHosp_bacterial,
                   a2_age_bacterial, a2_sex_bacterial, 
                   a2_subgroup_bacterial,
                   a2_total_bacterial, a2_bact_infection, a2_icu_mortality) %>%
            dplyr::mutate( site = site_id )
  
  rm( a2_age_all, a2_sex_all, a2_total_all, a2_age_bacterial, a2_sex_bacterial,a2_subgroup_bacterial,
      a2_lengthHosp_all, a2_lengthHosp_bacterial,
      a2_total_bacterial, a2_bact_infection, a2_icu_mortality )
  
  # put together all the sites a2 tables as a2_all_to_work      
  if( i ==1 ){
    a2_all_to_work <- a2_all
    rm(a2_all)
    }else{
      a2_all_to_work <- rbind( a2_all_to_work, a2_all)
      rm(a2_all)
    }
}
```

Remove from the environment all the variables except the dir.input, the files and the new created a2_all_to_work one. 
```{r}
rm(list=setdiff(ls(), c("a2_all_to_work", "files", "dir.input")))
```


##### A1 tables
```{r}
for( i in 1:length( files )){
  # files for each site
  site_id <- files[i]
  
  #load the RData file with all the a1 tables (aggregate counts per NPI period)
  load(paste0(dir.input, files[i], "/", "a1Tables.RData"))
  
  a1_total_all <-a1_total_all %>%
    dplyr::mutate( category_name = "total", type = "all" )
  
  a1_age_all <- a1_age_all %>%
    dplyr::mutate( category_name = "age", type = "all" )
  
  a1_sex_all <- a1_sex_all %>%
    dplyr::mutate( category_name = "sex", type = "all" )
  
  a1_lengthHosp_all <- a1_lengthHosp_all %>%
    dplyr::mutate( category_name = "lengthHospitalization", type = "all" )
  
  a1_season_all <- a1_season_all %>%
    dplyr::mutate( category_name = "season", type = "all" )
 
  a1_total_bacterial <-a1_total_bacterial %>%
    dplyr::mutate( category_name = "total", type = "bacterial" )
  
  a1_age_bacterial <- a1_age_bacterial %>%
    dplyr::mutate( category_name = "age", type = "bacterial" )
  
  a1_sex_bacterial <- a1_sex_bacterial %>%
    dplyr::mutate( category_name = "sex", type = "bacterial" )
  
  a1_lengthHosp_bacterial <- a1_lengthHosp_bacterial %>%
    dplyr::mutate( category_name = "lengthHospitalization", type = "bacterial" ) 
  
  a1_bact_infection <- a1_bact_infection %>%
    dplyr::mutate( category_name = "bacterialInfection", type = "bacterial" ) %>%
    dplyr::rename( category = disorder_group )
  
  a1_season_bacterial <- a1_season_bacterial %>%
    dplyr::mutate( category_name = "season", type = "bacterial" )
 
   a1_subgroup_bacterial <- a1_subgroups %>%
    dplyr::mutate( category_name = "subgroups", type = "bacterial" )
 
  
  ### add dead and ICU, in this case the category is a number so we will need to change
  ### it later to character to merge with the rest of tables
  a1_icu_all <- a1_icu_all %>%
    dplyr::mutate( category_name = "icu",type = "all" )
  
  a1_icu_bacterial <- a1_icu_bacterial %>%
    dplyr::mutate( category_name = "icu", type = "bacterial" )
  
  a1_mortality_all <- a1_mortality_all %>%
    dplyr::mutate( category_name = "dead", type = "all" )
  
  a1_mortality_bacterial <- a1_mortality_bacterial %>%
    dplyr::mutate( category_name = "dead", type = "bacterial" )
  
  a1_icu_mortality <- rbind( a1_icu_all, a1_icu_bacterial,
                             a1_mortality_all, a1_mortality_bacterial)
  
  # remove the original data frames
  rm(a1_icu_all, a1_icu_bacterial, a1_mortality_all, a1_mortality_bacterial)
  
  # change the type of the category column to combine with the rest of a1 tables
  a1_icu_mortality <- a1_icu_mortality %>%
    dplyr::mutate( category = as.character( category ))
  
  # add all the a1 tables together as a1_all and remove the original ones 
  # add a column with the site id
  a1_all <- rbind( a1_age_all, a1_sex_all, a1_total_all, 
                   a1_age_bacterial, a1_sex_bacterial, 
                   a1_lengthHosp_all, a1_lengthHosp_bacterial,
                   a1_season_bacterial, a1_season_all,
                   a1_subgroup_bacterial,
                   a1_total_bacterial, a1_bact_infection, a1_icu_mortality) %>%
            dplyr::mutate( site = site_id )
  
  rm( a1_age_all, a1_sex_all, a1_total_all, a1_age_bacterial, a1_sex_bacterial,a1_subgroup_bacterial,
      a1_lengthHosp_all, a1_lengthHosp_bacterial,
      a1_season_bacterial, a1_season_all,
      a1_total_bacterial, a1_bact_infection, a1_icu_mortality )
  
  # put together all the sites a1 tables as a1_all_to_work      
  if( i ==1 ){
    a1_all_to_work <- a1_all
    rm(a1_all)
    }else{
      a1_all_to_work <- rbind( a1_all_to_work, a1_all)
      rm(a1_all)
    }
}
```

Remove from the environment all the variables except the dir.input, the files, the a2_all_to_work and the new created a1_all_to_work one. 
```{r}
rm(list=setdiff(ls(), c("a2_all_to_work","a1_all_to_work", "files", "dir.input")))
ls()
```

Save the files generated. 
```{r}
save(a1_all_to_work, file = "../results/bacterial_aggregate_ouputs/a1_all_to_work.RData")
save(a2_all_to_work, file = "../results/bacterial_aggregate_ouputs/a2_all_to_work.RData")
```

