---
title: "Temporal analysis of bacterial Infection Over Time"
author: "Xiudi Li"
date: "2024-02-23"
output:
  word_document: default
  html_document: default
---

# Load libraries and set up the sites
```{r setup, include=FALSE}
rm(list=ls())

library(dplyr)
library(glmnet)
library(ggplot2)
library(gridExtra)

# "a2Tables.RData" and "a1Tables.RData" are stored in folders called "BCH", "CCMC", ..... 
sites = c("BCH","CCMC","GOSH","H12O","HCUV","MICHIGAN", "UPitt", "CCHMC")
```

````{r}
dat_combined = NULL
design.mats = vector("list",length(sites))
# allow a possible change in slope every gap=3 months
gap = 3 #use gap=2 as sensitivity analysis, change points identified are very consistent

#disease_subgroupToStudy <- "Skin"
disease_subgroupToStudy <- "Respiratory"
#disease_subgroupToStudy <- "IBI_extensive"
#disease_subgroupToStudy <- "IBI_simple"

for (kk in 1:length(sites)){
  load(paste0("../results/sites_outputs/", sites[kk],"/a2Tables.RData"))
  load(paste0("../results/sites_outputs/", sites[kk],"/a1Tables.RData"))
  
  # this line looks at total bacterial infection
  #dat = a2_total_bacterial
  # this line looks at a subgroup of bacterial infections
  dat = (a2_subgroups %>% filter(subgroup==disease_subgroupToStudy))[,-2]
  colnames(dat)[2] = "bacterial"
  dat = merge(dat,a2_total_all,all.y=TRUE)
  colnames(dat)[3] = "all"
  dat$bacterial[is.na(dat$bacterial)] = 0
  dat$frac = dat$bacterial/dat$all
  
  # creating the covariate matrix corresponding to age group
  age.frac = NULL
  for (age.group in c("childhood","infancy","middle childhood")){
    this_group = a2_age_all %>% filter(category == age.group) %>% arrange(timeToGroup)
    this_group = merge(dat,this_group,all.x=TRUE)
    this_group$count = ifelse(is.na(this_group$count),0,this_group$count)
    age.frac = cbind(age.frac,this_group$count/dat$all)
  }
  colnames(age.frac) = c("cFrac","iFrac","mFrac")
  dat = cbind(dat,age.frac)
  
  # parse date into month, relative month, and season
  dat$month = sapply(dat$timeToGroup,function(x){as.numeric(strsplit(as.character(x),split="-")[[1]][2])})
  dat$year = sapply(dat$timeToGroup,function(x){as.numeric(strsplit(as.character(x),split="-")[[1]][1])})
  dat$time = 12*(dat$year-2019) + dat$month
  print(c(sites[kk],max(dat$time),nrow(dat)))
  dat$season = rep(NA,nrow(dat))
  for (i in 1:nrow(dat)){
    if (dat$month[i] %in% c(1,2,12)){
      dat$season[i] = "winter"
    } else if (dat$month[i] %in% c(3,4,5)){
      dat$season[i] = "spring"
    } else if (dat$month[i] %in% c(6,7,8)){
      dat$season[i] = "summer"
    } else if (dat$month[i] %in% c(9,10,11)){
      dat$season[i] = "autumn"
    }
  }
  dat$site = rep(sites[kk],nrow(dat))
  dat = dat %>% filter(timeToGroup <= as.Date("2023-03-01"))
  #print(dim(dat))
  #print(range(dat$timeToGroup))
  if( sites[[kk]]== "CCHMC"){
    dat <- dat %>% filter( timeToGroup < as.Date("2022-04-01"))
  }
  dat_combined = rbind(dat_combined,dat)
  
  # creating the covariate matrix corresponding to season
  design.season = matrix(0,nrow(dat),3)
  seasons = c("summer","autumn","winter")
  for (j in 1:length(seasons)){
    design.season[,j] = 1*(dat$season == seasons[j])
  }
  colnames(design.season) = seasons
  
  # creating the covariate matrix corresponding to the linear spline basis
  nbasis = ceiling(51/gap)-1
  basis = matrix(0,nrow(dat),nbasis)
  for (i in 1:nbasis){
    basis[,i] = (dat$time - gap*i)*(dat$time > (gap*i))
  }
  design.mat = cbind(dat[,c("all","cFrac","iFrac","mFrac","time")],basis)
  colnames(design.mat)[-c(1:5)] = sapply(1:nbasis,function(i){paste0("time",gap*i)})
  design.mat = as.matrix(design.mat)
  design.mat = cbind(design.mat,design.season)
  #print(dim(design.mat))
  
  # each site has its own design matrix, store them in a list
  design.mats[[kk]] = design.mat
}

# creating the design matrix corresponding to site indicator
design.sites = matrix(0,nrow(dat_combined),length(sites)-1)
for (jj in 1:length(sites[-1])){
  design.sites[,jj] = 1*(dat_combined$site == sites[-1][jj])
}
colnames(design.sites) = sites[-1]

# combine all the pieces of design matrix
# the final combined design matrix is "design_combined"
design_combined = NULL
for (kk in 1:length(sites)){
  design_combined = rbind(design_combined,design.mats[[kk]])
}
design_combined = cbind(design_combined,design.sites)

```

We analyze data from all the sites. 
Let us visualized the counts in the 8 sites over time.

```{r echo=FALSE}
panelA <- ggplot(dat_combined) +
  geom_line(aes(x = timeToGroup, y = bacterial, colour = site)) + 
  labs(x="time",y="number of infection") +
  theme(legend.position = "none")
  
```

First, we fit a poisson GLM using total number of hospitalization as an offset, allowing change in slope every 3 months. That is,
\begin{equation}
 Y = \beta_0 + \sum_{j=1}^{16} \beta_j(t-t_j)_{+} + \beta_{age}* age + \beta_{season}* season + \alpha_{site}* site
\end{equation}
We fit the model using adaptive lasso, so that we achieve variable selection as well.

```{r echo=FALSE, warning=FALSE}
# -------------------------------------------------
# --- glm poisson model with total as offset ------
# -------------------------------------------------
set.seed(123)
mod.lasso.pois = cv.glmnet(x=design_combined[,-1], y=dat_combined$bacterial,family=poisson,offset=log(design_combined[,1]))
beta.init = as.vector(coef(mod.lasso.pois,s="lambda.1se"))[-1]
set.seed(123)
mod.lasso.pois = cv.glmnet(x=design_combined[,-1], y=dat_combined$bacterial,family=poisson,offset=log(design_combined[,1]),penalty.factor=1/abs(beta.init))
beta.alasso = coef(mod.lasso.pois,s="lambda.1se")
beta.alasso
pred.lasso = predict(mod.lasso.pois, s="lambda.1se",newx=design_combined[,-1],newoffset=log(design_combined[,1]),type="response")
dat_combined$pred = pred.lasso
```

Now let's visualize the fitted results in each site.

```{r echo = FALSE}
pred.lasso = predict(mod.lasso.pois, s="lambda.1se",newx=design_combined[,-1],newoffset=log(design_combined[,1]),type="response")
dat_combined$pred = pred.lasso
p1= ggplot(dat_combined %>% filter(site=="BCH")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("BCH")+labs(x="time",y="bacterial infection")
p2= ggplot(dat_combined %>% filter(site=="MICHIGAN")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("Michigan")+labs(x="time",y="bacterial infection")
p3= ggplot(dat_combined %>% filter(site=="HCUV")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("HCUV")+labs(x="time",y="bacterial infection")
p4= ggplot(dat_combined %>% filter(site=="H12O")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("H12O")+labs(x="time",y="bacterial infection")
grid.arrange(p1,p2,p3,p4,ncol=2)
p5= ggplot(dat_combined %>% filter(site=="CCMC")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("CCMC")+labs(x="time",y="bacterial infection")
grid.arrange(p1,p2,p3,p4,ncol=2)
p6= ggplot(dat_combined %>% filter(site=="GOSH")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("GOSH")+labs(x="time",y="bacterial infection")
p7= ggplot(dat_combined %>% filter(site=="UPitt")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("UPitt")+labs(x="time",y="bacterial infection")
p8= ggplot(dat_combined %>% filter(site=="CCHMC")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("CCHMC")+labs(x="time",y="bacterial infection")

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,ncol=4)
```

Next, based on the variable selection results. Let us re-fit the model using glm. 

```{r echo=FALSE}
# we now refit the model with variables selected by adaptive lasso
# this is to facilitate hypothesis testing
selected_var = rownames(beta.alasso)[which(beta.alasso != 0)][-1]
selected_var = union(selected_var, c("cFrac","iFrac","mFrac","summer","autumn","winter",
                 "CCHMC","CCMC","GOSH","H12O","HCUV","MICHIGAN","UPitt"))

# for Skin, IBI, do the following instead
#selected_var = c("cFrac","iFrac","mFrac","summer","autumn","winter",
#                 "CCHMC","CCMC","GOSH","H12O","HCUV","MICHIGAN","UPitt","time12","time24")

design_reduced = design_combined[,colnames(design_combined) %in% selected_var]
time_points = as.numeric(sapply(selected_var[sapply(selected_var,function(x){substr(x,1,4)}) == "time"],function(x){substr(x,5,nchar(x))}))
time_points
design_slope = data.frame(design_reduced)

# model with only slope change
dat_slope = cbind(dat_combined$bacterial,design_combined[,1],design_slope)
colnames(dat_slope)[1:2] = c("bacterial","all")
myformula = paste0("bacterial~offset(log(all))+",paste0(colnames(dat_slope)[-c(1:2)],collapse="+"))
mod_slope = glm(as.formula(myformula),family=poisson(),data=dat_slope)
summary(mod_slope)
```

And let's visualize the fit.

```{r echo=FALSE}
dat_combined$pred = predict(mod_slope,type="response")
p1= ggplot(dat_combined %>% filter(site=="BCH")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("BCH")+labs(x="time",y="bacterial infection")
p2= ggplot(dat_combined %>% filter(site=="MICHIGAN")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("Michigan")+labs(x="time",y="bacterial infection")
p3= ggplot(dat_combined %>% filter(site=="HCUV")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("HCUV")+labs(x="time",y="bacterial infection")
p4= ggplot(dat_combined %>% filter(site=="H12O")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("H12O")+labs(x="time",y="bacterial infection")
p5= ggplot(dat_combined %>% filter(site=="CCMC")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("CCMC")+labs(x="time",y="bacterial infection")
p6= ggplot(dat_combined %>% filter(site=="GOSH")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("GOSH")+labs(x="time",y="bacterial infection")
p7= ggplot(dat_combined %>% filter(site=="UPitt")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("UPitt")+labs(x="time",y="bacterial infection")
p8= ggplot(dat_combined %>% filter(site=="CCHMC")) + geom_col(aes(x=timeToGroup,y=bacterial))+geom_line(aes(x=timeToGroup,y=pred))+ggtitle("CCHMC")+labs(x="time",y="bacterial infection")

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,ncol=4)
```

# Now let's plot the fit using a standard population
```{r}
all_pred = NULL
for (kk in 1:length(sites)){
  standard_site = sites[kk]
  print(standard_site)
  if (standard_site == "BCH"){
    dat_standard_site = dat_slope[which(rowSums(dat_slope[,c("CCMC","GOSH","H12O","HCUV","MICHIGAN", "UPitt", "CCHMC")])==0),]
  } else {
    dat_standard_site = dat_slope[which(dat_slope[,standard_site]==1),]
  }
  dat_standard_site$all = mean(dat_standard_site$all)
  dat_standard_site$cFrac = mean(dat_standard_site$cFrac)
  dat_standard_site$iFrac = mean(dat_standard_site$iFrac)
  dat_standard_site$mFrac = mean(dat_standard_site$mFrac)
  dat_standard_site$winter = rep(0,nrow(dat_standard_site))
  
  pred_standard_site = predict(mod_slope,newdata=dat_standard_site)
  pred_standard_site = predict(mod_slope,newdata=dat_standard_site,type="response")
  this_pred = data.frame(time=dat_combined$time[dat_combined$site == standard_site],
                         pred = pred_standard_site,
                         site = rep(standard_site,length(pred_standard_site)))
  all_pred = rbind(all_pred,this_pred)
}

library(ggplot2)
all_pred$date = dat_combined$timeToGroup
panelC <- ggplot(data=all_pred,aes(x=date,y=pred,colour=site)) + geom_line()+ylab("predicted bacterial infection")+
  geom_vline(xintercept=as.Date("2020-01-01"),linetype=2,colour="blue")+geom_vline(xintercept=as.Date("2021-01-01"),linetype=2,colour="blue")+geom_vline(xintercept=as.Date("2022-09-01"),linetype=2,colour="blue")+geom_vline(xintercept=as.Date("2022-12-01"),linetype=2,colour="blue")+
  geom_vline(xintercept=as.Date("2020-03-01"),linetype=2)+geom_vline(xintercept=as.Date("2021-03-01"),linetype=2)

```

# Create final panel figure
```{r}
library(ggpu)

# ggarrange(panelB,
#           ggarrange(panelA, panelC, 
#                     ncol = 2, 
#                     labels = c("B", "C")), 
#           nrow = 2, 
#           labels = "A"
#           ) 

top_row = ggarrange(panelA, panelC, ncol = 2, labels = c("A", "C"))
bottom_row = ggarrange( panelB, ncol = 1, labels = c("B"))
final_plot = ggarrange(top_row, bottom_row, ncol = 1)

ggsave(filename = "../../panelFigure.png",
       plot = final_plot, 
       width = 10, 
       #height = 450, 
       dpi=700)

```

