---
title: "Bacterial infection count patients QC"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/4ceData/Input/4CEAllAdminChortsOct2023/"
obfuscation =  FALSE
#obfuscation = 3
dateFormat <- "%Y-%m-%d"
data_update_date <- "2023-10-01"
dir.output <- "/4ceData/bacterialInfection_outputs/"
age_estimation_variability <- FALSE
```

```{r message=FALSE, warning=FALSE}
#folder_4ce_files <- "/users/bha6ki/Documents/harvard_4CE_2023_tables/"
#obfuscation =  FALSE
#dateFormat <- "%Y-%m-%d"
#data_update_date <- "2023-03-01"
#dir.output <- "/users/bha6ki/Documents/Phase2.2.bacterial_infections-main/outputs/"
#age_estimation_variability <- FALSE
```


```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
source("../R/packagesInstallation.R")
# set up the theme plots
source("../R/theme_plots.R")
# get the session information
sessionInfo()
# load the local variables
source("../R/setUpVariableParameters.R")
# load the admission plot code
source("../R/admission_plot_count.R")


print(paste0( 'FourCePhase2.2bacterial Infection Code version: ', codeVersion ))
print('##########')
```

# Data loading
We will use as input the 2.2 data files. Specifically:
-   LocalPatientSummary
-   LocalPatientObservation
-   LocalPatientClinicalcourse

## Reading 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files without duplicating the data

source("../R/readInputFiles_fread.R")

filecheck <- checkInputFilesFread( path = folder_4ce_files, separator = ",", skip = 0, verbose = TRUE )

if(filecheck){
  demo_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientSummary.csv"), sep = ',',skip = 0, fill=TRUE )
  obs_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientObservations.csv"), sep = ',',skip = 0, fill=TRUE)
  clinical_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientClinicalCourse.csv"), sep = ',',skip = 0, fill=TRUE)
  


  ### Extract the patient summary and observation information. 
  # arrange in a consistent way to make sure we are not assuming the order of some sites
  # select only the study period
  demo_raw <- demo_raw %>% 
    dplyr::mutate( age = trunc( age ) ) %>%
    dplyr::filter(cohort == "AllAdm", 
                  age >= 0,
                  as.Date(admission_date, format = dateFormat) >= ymd('2019-01-01'),
                  as.Date(admission_date, format = dateFormat) <= ymd('2023-03-31')) %>%
    dplyr::arrange(patient_num, as.Date(admission_date, format = dateFormat), days_since_admission)
  obs_raw <- obs_raw %>% 
    dplyr::filter(cohort == "AllAdm") %>%
    dplyr::arrange(patient_num, days_since_admission)
  clinical_raw <- clinical_raw %>% 
    dplyr::filter(cohort == "AllAdm",
                  as.Date(calendar_date, format = dateFormat) >= ymd('2019-01-01'),
                  as.Date(calendar_date, format = dateFormat) <= ymd('2023-03-31')) %>%
    dplyr::arrange(patient_num, as.Date(calendar_date, format = dateFormat), days_since_admission)
  
  print('Input files successfully read, patient data extracted')
  
}else{
  print('Error in reading files')
}

#remove potential duplicates
clinical_raw <- clinical_raw %>% unique()
demo_raw <- demo_raw %>% unique()
obs_raw <- obs_raw %>% unique()
```
## CHECK1: how many patients are present in the 3 input files?
```{r}
demo_raw_patients <- unique( demo_raw$patient_num )
print(length(demo_raw_patients))

clinical_raw_patients <- unique( clinical_raw$patient_num )
print(length(clinical_raw_patients))

obs_raw_patients <- unique( obs_raw$patient_num )
print(length(obs_raw_patients))

common_patients <- clinical_raw_patients[ clinical_raw_patients %in% obs_raw_patients ]
common_patients <- common_patients[ common_patients %in% demo_raw_patients ]
print(length(common_patients))
```



## Select the ICD codes of interest
```{r}
#### ICD codes bacterial conditions ####
print('Reading ICD bacterial codes')
icdCodes <- read.csv("../public-data/ICD_bacterial5.csv", header = TRUE, colClasses = "character") %>%
   dplyr::mutate( concept_code = gsub("\\.", "",ICD10_Code)) %>%
   dplyr::select( disorder_group = syndrome, ICD10_Code, concept_code, description, 
                  syndrome, IBI_simple, IBI_extensive, Respiratory, Skin, bacterial_syndrome )

```

# Data-management

## Filtering on "inpatient" status

```{r}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- dplyr::filter(clinical_raw, in_hospital == 1)
```

## Calculating duration of individual hospitalizations

- Adding columns with hospitalization number and hospitalization length to
the clinical raw file
- Calculate 'admit date': date of admission for each hospitalization

```{r}
source("../R/count_sequences_hospitalisation.R")
clinical_raw <- count_hosp(clinical_raw)

print('Hospitalization sequences have been counted.')
```

## CHECK2: how many inpatients with ICD codes since admission are present in the 3 input files?
```{r}
clinical_raw_patients_filtered <- unique(clinical_raw$patient_num)
print(length(clinical_raw_patients_filtered))

obs_raw_patients_filtered <- obs_raw %>%
  dplyr::filter( trimws(concept_type) == "DIAG-ICD10" ) %>%
  dplyr::filter( days_since_admission >= 0 ) %>%
  dplyr::select( patient_num ) %>%
  unique()

obs_raw_patients_filtered <- obs_raw_patients_filtered$patient_num
print(length(obs_raw_patients))

common_patients_filtered <- clinical_raw_patients_filtered[ clinical_raw_patients_filtered %in% obs_raw_patients_filtered ]
print(length(common_patients_filtered))
```

```{r}
clinical_raw <- clinical_raw %>%
  dplyr::mutate( month_hosp = as.Date(admit_date, format = dateFormat), 
                 month_hosp = floor_date(ymd(month_hosp), "month")) %>%
  dplyr::ungroup()

```

## CHECK3: how many hospitalized patients per month
```{r}
monthly_hosp <- clinical_raw %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p1 <- admission_plot_count(df = monthly_hosp, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p1
```

## CHECK4: how many hospitalized patients per month that are in obs_raw
```{r}
obs_raw_pat <- unique( obs_raw$patient_num)

monthly_hosp_in_obs <- clinical_raw %>%
  dplyr::filter( patient_num %in% obs_raw_pat ) %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p2 <- admission_plot_count(df = monthly_hosp_in_obs, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p2
```


## CHECK5: how many hospitalized patients per month that are in obs_raw and have ICD codes
```{r}
obs_raw_inpatients_icd10 <- obs_raw %>%
  dplyr::filter( patient_num %in% clinical_raw$patient_num ) %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10") 

obs_raw_inpatients_icd10 <- unique( obs_raw_inpatients_icd10$patient_num)

monthly_hosp_in_obs_icd10 <- clinical_raw %>%
  dplyr::ungroup() %>%
  dplyr::filter( patient_num %in% obs_raw_inpatients_icd10 ) %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p3 <- admission_plot_count(df = monthly_hosp_in_obs_icd10, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p3

```


## CHECK6: how many hospitalized patients per month that are in obs_raw and have ICD codes after admission
```{r}
obs_raw_inpatients_icd10_aa <- obs_raw %>%
  dplyr::filter( patient_num %in% clinical_raw$patient_num ) %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10")  %>%
  dplyr::filter( days_since_admission >= 0 )

obs_raw_inpatients_icd10_aa <- unique(obs_raw_inpatients_icd10_aa$patient_num )

monthly_hosp_in_obs_icd10_aa <- clinical_raw %>%
  dplyr::ungroup() %>%
  dplyr::filter( patient_num %in% obs_raw_inpatients_icd10_aa) %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p4 <- admission_plot_count(df = monthly_hosp_in_obs_icd10_aa, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p4

```

## CHECK7: extract some patients that are lost when filtering by ICD10 code in april
```{r}
april_patients <- clinical_raw %>%
  dplyr::ungroup() %>%
  dplyr::filter( month_hosp == "2022-04-01")

print(length(unique(april_patients$patient_num)))

april_patients_in_obs_raw <- april_patients %>%
    dplyr::filter( patient_num %in% obs_raw$patient_num )

print(length(unique(april_patients_in_obs_raw$patient_num)))

april_patients_with_icd10_codes <- obs_raw %>%
    dplyr::filter( patient_num %in% april_patients$patient_num ) %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10") 

print(length(unique(april_patients_with_icd10_codes$patient_num)))

april_patients_with_icd10_codes_aa <- obs_raw %>%
    dplyr::filter( patient_num %in% april_patients$patient_num ) %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10") %>%
  dplyr::filter( days_since_admission >= 0 )

print(length(unique(april_patients_with_icd10_codes_aa$patient_num)))

### save the missing patients to check if they have other info in obs_raw
missing_patients_april <- april_patients[ ! april_patients$patient_num %in% april_patients_with_icd10_codes_aa$patient_num,]

explore_missing_patients_april <- obs_raw %>% 
  dplyr::filter( patient_num %in% missing_patients_april$patient_num ) %>%
  dplyr::left_join( april_patients, by = c("patient_num", "days_since_admission"))
summary(explore_missing_patients_april$days_since_admission)
summary(as.factor(explore_missing_patients_april$concept_type))
```


## Calculating age at hospitalization time

```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
# removing the dots from the ICD codes (if any)
obs_raw <- as.data.frame( obs_raw )
demo_raw <- as.data.frame( demo_raw )
clinical_raw <- as.data.frame( clinical_raw )

obs_raw_demo <- left_join(obs_raw, demo_raw[c("patient_num", "age", "dead", "death_date")], by = c("patient_num")) %>%
  dplyr::mutate( concept_code = gsub("\\.", "",concept_code )) 
```

## CHECK8: after merging the obs_raw with demo_raw
```{r}
obs_raw_inpatients_icd10_demo <- obs_raw_demo %>%
  dplyr::filter( patient_num %in% clinical_raw$patient_num ) %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10")  %>%
  dplyr::filter( days_since_admission >= 0 )


obs_raw_inpatients_icd10_demo <- unique(obs_raw_inpatients_icd10_demo$patient_num )

monthly_hosp_in_obs_icd10_aa_demo <- clinical_raw %>%
  dplyr::ungroup() %>%
  dplyr::filter( patient_num %in% obs_raw_inpatients_icd10_demo) %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p5 <- admission_plot_count(df = monthly_hosp_in_obs_icd10_aa_demo, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p5
```

## CHECK9:  after merging the obs_raw with demo_raw and clinical_raw
```{r}
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age", "dead", "death_date")], by = c("patient_num")) %>%
  dplyr::mutate( concept_code = gsub("\\.", "",concept_code )) %>%
  dplyr::inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date", "month_hosp")], 
            by = c("patient_num", "days_since_admission"))
```

```{r}
obs_raw_inpatients_icd10_demo_clin <- obs_raw_age %>%
  dplyr::filter( trimws( concept_type) == "DIAG-ICD10")  %>%
  dplyr::filter( days_since_admission >= 0 )

monthly_hosp_in_obs_icd10_aa_demo_clin <- obs_raw_inpatients_icd10_demo_clin %>%
  dplyr::ungroup() %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p6 <- admission_plot_count(df = monthly_hosp_in_obs_icd10_aa_demo, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p6
```


````{r}
# estimate age at each diagnosis
if( age_estimation_variability == FALSE ){
  
  obs_raw_age <- obs_raw_age %>%
    dplyr::mutate( age_time_diagnosis = age - floor(as.numeric(as.Date(data_update_date, format = dateFormat) - as.Date(calendar_date, format = dateFormat)) / 365.25)) %>%
    dplyr::select( -dead, -death_date )
  
}

print('Patient age calculated')
```

## CHECK10: make sure that the concept_type values for ICD10 look all the same
```{r}
concept_types <- unique( obs_raw_age$concept_type )
concept_types
```

## Check 11: counts after filtering by DIAG-ICD10
```{r}
obs_filtered_inclusion_criteria_icd10 <- obs_raw_age %>%
    dplyr::filter(concept_type == 'DIAG-ICD10')

to_check <- obs_filtered_inclusion_criteria_icd10 %>%
  dplyr::ungroup() %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p7 <- admission_plot_count(df = to_check, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p7
```


## Check 12: counts after filtering by age

```{r}
obs_filtered_inclusion_criteria <- obs_raw_age %>% 
  dplyr::filter(concept_type == 'DIAG-ICD10',
                days_since_admission >= 0, 
                age_time_diagnosis < 18 )

to_check_age <- obs_filtered_inclusion_criteria %>%
  dplyr::ungroup() %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p8 <- admission_plot_count(df = to_check_age, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p8
```
## Check 13: add a estimation of patients per month and age
```{r}
counts_month_age <- obs_filtered_inclusion_criteria %>%
  dplyr::mutate( year = floor_date(ymd(month_hosp), "year")) %>%
  dplyr::ungroup() %>%
  dplyr::group_by( age, month_hosp, year ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

b1 <- ggplot( data = counts_month_age, aes( x= month_hosp, y=n, fill="age")) +
  geom_bar( stat="identity") +
  facet_wrap( ~ age )

b1
```

## Check 14: get counts of length of hospitalization per month
```{r}
if (!time_period %in% c('day', 'week', 'month')) {
  print(paste0('unidentified time period: ', time_period))
  print(paste0('possible time periods: ', c('day', 'week', 'month')))
} else {
  print(paste0('time period: ', time_period))
}

# base the time period on admit date, not calendar date
clinical_filtered  <- clinical_raw %>%
  dplyr::filter( length_hospitalization >= min_hosp_days, 
          cohort  %in% c("AllAdm")) %>%
  dplyr::mutate( calendar_date = as.Date( calendar_date, format = dateFormat),
                 admit_date = as.Date( admit_date, format = dateFormat),
                 time_p = as.Date(cut( admit_date, breaks = time_period)),
                 period = ifelse( admit_date <= pre_NPI, "pre-NPI", 
                                  ifelse( admit_date > full_NPI, "partial-NPI", "full-NPI")))


to_check_lengthHosp <- clinical_filtered %>%
  dplyr::ungroup() %>%
  dplyr::group_by( month_hosp ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

p9 <- admission_plot_count(df = to_check_lengthHosp, 
                     x = "month_hosp",
                     y = "n", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
p9


counts_month_lenght_hosp <- clinical_filtered %>%
  dplyr::mutate( year = floor_date(ymd(month_hosp), "year")) %>%
  dplyr::ungroup() %>%
  dplyr::group_by( length_hospitalization, month_hosp, year ) %>%
  dplyr::summarize( n = n_distinct( patient_num ))

summary(counts_month_lenght_hosp$length_hospitalization)

b2 <- counts_month_lenght_hosp %>%
  dplyr::filter( length_hospitalization < 21 ) %>%
  ggplot(  aes( x= month_hosp, y=n, fill="length_hospitalization")) +
  geom_bar( stat="identity") +
  facet_wrap( ~ length_hospitalization )

b2
```



```{r}
# we create a hospitalization id and select only the columns of interest
clinical_filtered <- clinical_filtered %>%
  dplyr::mutate( hospitalization_id = paste0( patient_num, "-", n_hospitalization )) %>% # create id to each hospitaliz.
  dplyr::select( patient_num, days_since_admission, in_hospital, in_icu, dead, length_hospitalization, n_hospitalization, time_p, period, hospitalization_id ) %>%
  unique()
  

# we filter the demographic data by patients that meet the inclusion criteria
# from all the information there we select the patient id, the time patient die and sex
demographic_filtered_inclusion_criteria <- demo_raw %>%
  filter( patient_num %in% obs_filtered_inclusion_criteria$patient_num ) %>%
  select( patient_num, death_date, sex) %>%
  unique()
```

## CHECK15: how many inpatients with ICD codes since admission in the age range are present in the 3 input files?
```{r}
clinical_filtered_patients <- unique(clinical_filtered$patient_num)
print(length(clinical_filtered_patients))

obs_filtered_inclusion_criteria_patients <- unique( obs_filtered_inclusion_criteria$patient_num )
print(length(obs_filtered_inclusion_criteria_patients))

common_patients_filtered_2 <- clinical_filtered_patients[ clinical_filtered_patients %in% obs_filtered_inclusion_criteria_patients ]
print(length(common_patients_filtered_2))
```


## Put all the data together in a unique table

We create a table with all the data together. 
```{r}
all_codes <- obs_filtered_inclusion_criteria %>% 
   dplyr::inner_join( clinical_filtered, 
                     by = c("patient_num", "days_since_admission")) %>% 
   dplyr::inner_join(demographic_filtered_inclusion_criteria, 
                     by = "patient_num")
```

# Monthly hospitalizations

## Total (counts)
Estimate the number of monthly hospitalizations. 
For this estimate, done for QC purposes, we do not filter by any condition. 

```{r}
monthly_hospitalizations <- all_codes %>%
  dplyr::filter( time_p < end_date_plots, 
                 time_p >= start_date_plots ) %>%
  unique() %>%
  dplyr::group_by( time_p ) %>%
  dplyr::summarise(count_total =  ifelse(n_distinct(hospitalization_id) > obfuscation | isFALSE(obfuscation),
                               n_distinct(hospitalization_id), 
                               0.5)) %>%
 dplyr::mutate( period = ifelse( time_p <= pre_NPI,
                           "pre-NPI", ifelse( time_p > full_NPI, "partial-NPI", "full-NPI")))
nh <- admission_plot_count(df = monthly_hospitalizations, 
                     x = "time_p",
                     y = "count_total", 
                     title = paste0("Total hospitalizations (per ", time_period,")"), 
                     y_title = "Hospitalization Count")
nh
```
